---
title: "Alpha/Beta Diversities"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
---

```{r setup, warning=FALSE, message=FALSE}
# rm(list = ls())

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = FALSE, comment = NA, 
                      fig.width = 6.25, fig.height = 5)

library(openxlsx)
library(tidyverse) 
library(CLME)
library(ggprism)
library(ggsci)
library(ggpubr)
library(RColorBrewer)
library(microbiome)
library(jtools)
library(vegan)
library(rstatix)
library(RVAideMemoire)
```

```{r helper}
plot_clme = function(clme_obj, group, y_min, y_max, p_gap, ann_pos, decreasing = FALSE, ...) {
  overall_p = clme_obj$p.value
  ind_p = clme_obj$p.value.ind
  est_mean = clme_obj$theta[group]
  est_se = sqrt(diag(clme_obj$cov.theta))[group]
  
  df_fig = data.frame(x = group, y = est_mean, err = est_se)
  if (decreasing) {
    df_p = data.frame(group1 = group[seq_len(length(group) - 1)],
                      group2 = group[-1],
                      x = group[-1],
                      label = paste0("p = ", round(ind_p, 3)),
                      y.position = seq.int(from = ifelse(est_mean[1] > min(est_mean), 
                                                         est_mean[1] + p_gap, 
                                                         est_mean[1] + 2 * p_gap),
                                           to = min(est_mean) + p_gap, 
                                           length.out = length(group) - 1))
  } else {
    df_p = data.frame(group1 = group[seq_len(length(group) - 1)],
                      group2 = group[-1],
                      x = group[-1],
                      label = paste0("p = ", round(ind_p, 3)),
                      y.position = seq.int(from = est_mean[2] + p_gap, 
                                           to = ifelse(est_mean[2] < max(est_mean), 
                                                       max(est_mean) + p_gap, 
                                                       max(est_mean) + 2 * p_gap), 
                                           length.out = length(group) - 1))
  }
  
  df_ann = data.frame(x = group[1], y = ann_pos,
                      fill = "white",
                      label = paste0("Trend p-value = ", round(overall_p, 3)))
  
  fig = df_fig %>%
    ggplot(aes(x = x, y = y)) +
    geom_bar(stat = "identity", color = "black", aes(fill = x)) + 
    geom_errorbar(aes(ymin = y - 1.96 * err, ymax = y + 1.96 * err), 
                  width = .2, position = position_dodge(.9)) +
    add_pvalue(df_p,
               xmin = "group1",
               xmax = "group2",
               label = "label",
               y.position = "y.position",
               remove.bracket = FALSE, 
               ...) +
    geom_label(data = df_ann, aes(x = x, y = y, label = label), 
               size = 4, vjust = -0.5, hjust = 0, color = "black") +
    ylim(y_min, y_max) +
    theme_bw()
  return(fig)
}
```

```{r data}
# OTU table
otu_table = read_tsv("../data/otu-table.tsv")
otu_id = otu_table$`#OTU ID`
otu_table = data.frame(otu_table[, -1], check.names = FALSE, row.names = otu_id)

# Taxonomy table
tax = read_tsv("../data/taxonomy.tsv")
otu_id = tax$`Feature ID`
tax = data.frame(tax[, - c(1, 3)], row.names = otu_id)
tax = tax %>% 
  separate(col = Taxon, 
           into = c("Kingdom", "Phylum", "Class", "Order", 
                    "Family", "Genus", "Species"),
           sep = ";") %>%
  rowwise() %>%
  dplyr::mutate_all(function(x) strsplit(x, "__")[[1]][2]) %>%
  mutate(Species = ifelse(!is.na(Species) & !is.na(Genus),
                          paste(ifelse(strsplit(Genus, "")[[1]][1] == "[",
                                       strsplit(Genus, "")[[1]][2],
                                       strsplit(Genus, "")[[1]][1]), Species, sep = "."),
                          NA)) %>%
  ungroup()
tax = as.matrix(tax)
rownames(tax) = otu_id
tax[tax == ""] = NA

# Tree
tree = read_tree("../data/tree.nwk")

# Meta data
meta_data = read_csv("../data/df_v1.csv")
meta_data$status = factor(meta_data$status, levels = c("nc", "sc"))
meta_data$time2aids = factor(meta_data$time2aids,
                             levels = c("never", "> 10 yrs",
                                        "5 - 10 yrs", "< 5 yrs"))

# Phyloseq object
OTU = otu_table(otu_table, taxa_are_rows = TRUE)
META = sample_data(meta_data)
sample_names(META) = meta_data$sampleid
TAX = tax_table(tax)
otu_data = phyloseq(OTU, TAX, META, tree)
genus_data = aggregate_taxa(otu_data, "Genus")
species_data = aggregate_taxa(otu_data, "Species")

genus_rarefied = rarefy_even_depth(genus_data, rngseed = 1, 
                                   sample.size = 0.9 * min(sample_sums(genus_data)), 
                                   replace = FALSE)
species_rarefied = rarefy_even_depth(species_data, rngseed = 1, 
                                     sample.size = 0.9 * min(sample_sums(species_data)), 
                                     replace = FALSE)
```

# Primary grouping

## Alpha diversities

1. Observed species: richness measure, which returns observed richness.

2. Shannon index: diversity measure, which takes into account richness, divergence and evenness.

3. P-value is obtained by linear regressions adjusting for abx usage.

### Species level {.tabset}

```{r}
d_alpha = alpha(species_rarefied, index = c("observed", "diversity_shannon"))
df_alpha = data.frame(sampleid = meta(species_rarefied)$sampleid,
                      subjid = meta(species_rarefied)$subjid,
                      group = meta(species_rarefied)$group1,
                      abx_use = meta(species_rarefied)$abx_use,
                      Observed = d_alpha$observed,
                      Shannon = d_alpha$diversity_shannon,
                      check.names = FALSE) %>%
  filter(group != "missing")
df_alpha$group = recode(df_alpha$group, 
                        `g1` = "G1", `g2` = "G2",
                        `g3` = "G3", `g4` = "G4")

df_fig = df_alpha %>% 
  gather(key = "measure", value = "value", Observed:Shannon)
df_fig$measure = factor(df_fig$measure, 
                        levels = c("Observed", "Shannon"))
df_fig$group = factor(df_fig$group, levels = c("G1", "G2", "G3", "G4"))
stat_test1 = df_fig %>%
  filter(measure == "Observed") %>%
  lm(formula = value ~ group + abx_use)
stat_test2 = df_fig %>%
  filter(measure == "Shannon") %>%
  lm(formula = value ~ group + abx_use)
```

#### Observed species

```{r}
summ(stat_test1)
```

#### Shannon index

```{r}
summ(stat_test2)
```

#### Box plot

```{r, fig.width=10, fig.height=5}
df_stat = data.frame(p = round(summary(stat_test1)$coef[2:4, "Pr(>|t|)"], 2)) %>%
  mutate(x = "group", y = "value", measure = "Observed", 
         group1 = "G1", group2 = c("G2", "G3", "G4"),
         y.position = max(stat_test1$model$value) + seq(0, 2, 1)) %>%
  bind_rows(
    data.frame(p = round(summary(stat_test2)$coef[2:4, "Pr(>|t|)"], 2)) %>%
      mutate(x = "group", y = "value", measure = "Shannon", 
             group1 = "G1", group2 = c("G2", "G3", "G4"),
             y.position = max(stat_test2$model$value) + seq(0, 0.4, 0.2))
  ) 

p = df_fig %>%
  ggboxplot(x = "group", y = "value", color = "group",
            add = "jitter", 
            xlab = FALSE, ylab = FALSE, title = FALSE) +
  stat_pvalue_manual(df_stat, label = "p", tip.length = 0.01) +
  scale_color_npg(name = "Group") +
  facet_wrap(.~measure, scale = "free", nrow = 1) + 
  theme_bw() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5))
p
```

#### Monotonically increasing trend test

```{r, fig.width=12, fig.height=5}
df_alpha$group = factor(df_alpha$group, levels = c("G1", "G2", "G3", "G4"), ordered = TRUE)
cons = list(order = "simple", decreasing = FALSE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_increase_species = plot_clme(summ_fit1, group = c("G1", "G2", "G3", "G4"),
                                  y_min = 0, y_max = 7, p_gap = 0.7, ann_pos = 6.5)
fig1_increase_species = fig1_increase_species +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_increase_species = plot_clme(summ_fit2, group = c("G1", "G2", "G3", "G4"),
                                  y_min = 0, y_max = 1.4, p_gap = 0.2, ann_pos = 1.3)
fig2_increase_species = fig2_increase_species +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

#### Monotonically decreasing trend test

```{r, fig.width=12, fig.height=5}
cons = list(order = "simple", decreasing = TRUE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_decrease_species = plot_clme(summ_fit1, group = c("G1", "G2", "G3", "G4"),
                                  y_min = 0, y_max = 7, p_gap = 0.8, ann_pos = 6.5, decreasing = TRUE)
fig1_decrease_species = fig1_decrease_species +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_decrease_species = plot_clme(summ_fit2, group = c("G1", "G2", "G3", "G4"),
                                  y_min = 0, y_max = 1.4, p_gap = 0.1, ann_pos = 1.3, decreasing = TRUE)
fig2_decrease_species = fig2_decrease_species +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

### Genus level {.tabset}

```{r}
d_alpha = alpha(genus_rarefied, index = c("observed", "diversity_shannon"))
df_alpha = data.frame(sampleid = meta(genus_rarefied)$sampleid,
                      subjid = meta(genus_rarefied)$subjid,
                      group = meta(genus_rarefied)$group1,
                      abx_use = meta(genus_rarefied)$abx_use,
                      Observed = d_alpha$observed,
                      Shannon = d_alpha$diversity_shannon,
                      check.names = FALSE) %>%
  filter(group != "missing")
df_alpha$group = recode(df_alpha$group, 
                        `g1` = "G1", `g2` = "G2",
                        `g3` = "G3", `g4` = "G4")

df_fig = df_alpha %>% 
  gather(key = "measure", value = "value", Observed:Shannon)
df_fig$group = factor(df_fig$group, levels = c("G1", "G2", "G3", "G4"))
df_fig$measure = factor(df_fig$measure, 
                        levels = c("Observed", "Shannon"))
stat_test1 = df_fig %>%
  filter(measure == "Observed") %>%
  lm(formula = value ~ group + abx_use)
stat_test2 = df_fig %>%
  filter(measure == "Shannon") %>%
  lm(formula = value ~ group + abx_use)
```

#### Observed genus

```{r}
summ(stat_test1)
```

#### Shannon index

```{r}
summ(stat_test2)
```

#### Box plot

```{r, fig.width=10, fig.height=5}
df_stat = data.frame(p = round(summary(stat_test1)$coef[2:4, "Pr(>|t|)"], 2)) %>%
  mutate(x = "group", y = "value", measure = "Observed", 
         group1 = "G1", group2 = c("G2", "G3", "G4"),
         y.position = max(stat_test1$model$value) + seq(0, 2, 1)) %>%
  bind_rows(
    data.frame(p = round(summary(stat_test2)$coef[2:4, "Pr(>|t|)"], 2)) %>%
      mutate(x = "group", y = "value", measure = "Shannon", 
             group1 = "G1", group2 = c("G2", "G3", "G4"),
             y.position = max(stat_test2$model$value) + seq(0, 0.4, 0.2))
  ) 

p = df_fig %>%
  ggboxplot(x = "group", y = "value", color = "group",
            add = "jitter", 
            xlab = FALSE, ylab = FALSE, title = FALSE) +
  stat_pvalue_manual(df_stat, label = "p", tip.length = 0.01) +
  scale_color_npg(name = "Group") +
  facet_wrap(.~measure, scale = "free", nrow = 1) + 
  theme_bw() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5))
p
```

#### Monotonically increasing trend test

```{r, fig.width=12, fig.height=5}
df_alpha$group = factor(df_alpha$group, levels = c("G1", "G2", "G3", "G4"), ordered = TRUE)
cons = list(order = "simple", decreasing = FALSE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_increase_genus = plot_clme(summ_fit1, group = c("G1", "G2", "G3", "G4"),
                                y_min = 0, y_max = 12, p_gap = 0.8, ann_pos = 11)
fig1_increase_genus = fig1_increase_genus +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_increase_genus = plot_clme(summ_fit2, group = c("G1", "G2", "G3", "G4"),
                                y_min = 0, y_max = 2.5, p_gap = 0.2, ann_pos = 2.2)
fig2_increase_genus = fig2_increase_genus +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

#### Monotonically decreasing trend test

```{r, fig.width=12, fig.height=5}
cons = list(order = "simple", decreasing = TRUE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_decrease_genus = plot_clme(summ_fit1, group = c("G1", "G2", "G3", "G4"),
                                y_min = 0, y_max = 12.5, p_gap = 1.1, ann_pos = 11.5, decreasing = TRUE)
fig1_decrease_genus = fig1_decrease_genus +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_decrease_genus = plot_clme(summ_fit2, group = c("G1", "G2", "G3", "G4"),
                                y_min = 0, y_max = 2.5, p_gap = 0.2, ann_pos = 2.2, decreasing = TRUE)
fig2_decrease_genus = fig2_decrease_genus +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

## Beta diversity {.tabset}

1. P-values are obtained by PERMANOVA

2. Dissimilarity measure: Bray-Curtis

### Species level

```{r, fig.width=8, fig.height=6}
species_rarefied = subset_samples(species_rarefied, group1 != "missing")

dis = phyloseq::distance(species_rarefied, method = "bray")
mds  = ordinate(species_rarefied, "MDS", distance = dis)

# PCoA plot
fig_beta_species = plot_ordination(species_rarefied, 
                      mds, color = "group1") +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(geom = "polygon", type = "t", level = 0.8,
               aes(fill = group1), alpha = 0.25) +
  scale_color_npg(name = NULL, labels = c("G1", "G2", "G3", "G4")) +
  scale_fill_npg(name = NULL, labels = c("G1", "G2", "G3", "G4")) +
  theme_bw()

# P-values
groups = meta(species_rarefied)$group1
set.seed(123)
stat_test = pairwise.perm.manova(resp = dis, fact = groups, nperm = 999, p.method = "none")
df_p = data.frame(X = rep("G1", 3),
                  Y = c("G2", "G3", "G4"),
                  p = round(stat_test$p.value[, "g1"], 2),
                  q = round(p.adjust(stat_test$p.value[, "g1"], method = "BH"), 2))

tab_beta_species = ggtexttable(df_p, rows = NULL, theme = ttheme("light"))
```

### Genus level

```{r, fig.width=8, fig.height=6}
genus_rarefied = subset_samples(genus_rarefied, group1 != "missing")

dis = phyloseq::distance(genus_rarefied, method = "bray")
mds  = ordinate(genus_rarefied, "MDS", distance = dis)

# PCoA plot
fig_beta_genus = plot_ordination(genus_rarefied, 
                      mds, color = "group1") +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(geom = "polygon", type = "t", level = 0.8,
               aes(fill = group1), alpha = 0.25) +
  scale_color_npg(name = NULL, labels = c("G1", "G2", "G3", "G4")) +
  scale_fill_npg(name = NULL, labels = c("G1", "G2", "G3", "G4")) +
  theme_bw()

# P-values
groups = meta(genus_rarefied)$group1
set.seed(123)
stat_test = pairwise.perm.manova(dis, groups, nperm = 999, p.method = "none")
df_p = data.frame(X = rep("G1", 3),
                  Y = c("G2", "G3", "G4"),
                  p = round(stat_test$p.value[, "g1"], 2),
                  q = round(p.adjust(stat_test$p.value[, "g1"], method = "BH"), 2))

tab_beta_genus = ggtexttable(df_p, rows = NULL, theme = ttheme("light"))
```

# Secondary grouping

## Alpha diversities

1. Observed species: richness measure, which returns observed richness.

2. Shannon index: diversity measure, which takes into account richness, divergence and evenness.

3. P-value is obtained by linear regressions adjusting for abx usage.

### Species level {.tabset}

```{r}
d_alpha = alpha(species_rarefied, index = c("observed", "diversity_shannon"))
df_alpha = data.frame(sampleid = meta(species_rarefied)$sampleid,
                      subjid = meta(species_rarefied)$subjid,
                      group = meta(species_rarefied)$group2,
                      abx_use = meta(species_rarefied)$abx_use,
                      Observed = d_alpha$observed,
                      Shannon = d_alpha$diversity_shannon,
                      check.names = FALSE) %>%
  filter(group != "missing")
df_alpha$group = recode(df_alpha$group, 
                        `g1` = "G1", `g2` = "G2",
                        `g3` = "G3", `g4` = "G4", `g5` = "G5")

df_fig = df_alpha %>% 
  gather(key = "measure", value = "value", Observed:Shannon)
df_fig$measure = factor(df_fig$measure, 
                        levels = c("Observed", "Shannon"))
df_fig$group = factor(df_fig$group, levels = c("G1", "G2", "G3", "G4", "G5"))
stat_test1 = df_fig %>%
  filter(measure == "Observed") %>%
  lm(formula = value ~ group + abx_use)
stat_test2 = df_fig %>%
  filter(measure == "Shannon") %>%
  lm(formula = value ~ group + abx_use)
```

#### Observed species

```{r}
summ(stat_test1)
```

#### Shannon index

```{r}
summ(stat_test2)
```

#### Box plot

```{r, fig.width=10, fig.height=5}
df_stat = data.frame(p = round(summary(stat_test1)$coef[2:5, "Pr(>|t|)"], 2)) %>%
  mutate(x = "group", y = "value", measure = "Observed", 
         group1 = "G1", group2 = c("G2", "G3", "G4", "G5"),
         y.position = max(stat_test1$model$value) + seq(0, 3, 1)) %>%
  bind_rows(
    data.frame(p = round(summary(stat_test2)$coef[2:5, "Pr(>|t|)"], 2)) %>%
      mutate(x = "group", y = "value", measure = "Shannon", 
             group1 = "G1", group2 = c("G2", "G3", "G4", "G5"),
             y.position = max(stat_test2$model$value) + seq(0, 0.6, 0.2))
  ) 

p = df_fig %>%
  ggboxplot(x = "group", y = "value", color = "group",
            add = "jitter", 
            xlab = FALSE, ylab = FALSE, title = FALSE) +
  stat_pvalue_manual(df_stat, label = "p", tip.length = 0.01) +
  scale_color_npg(name = "Group") +
  facet_wrap(.~measure, scale = "free", nrow = 1) + 
  theme_bw() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5))
p
```

#### Monotonically increasing trend test

```{r, fig.width=12, fig.height=5}
df_alpha$group = factor(df_alpha$group, levels = c("G1", "G2", "G3", "G4", "G5"), ordered = TRUE)
cons = list(order = "simple", decreasing = FALSE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_increase_species_secondary = plot_clme(summ_fit1, 
                                            group = c("G1", "G2", "G3", "G4", "G5"),
                                            y_min = 0, y_max = 7, p_gap = 0.7, ann_pos = 6.5)
fig1_increase_species_secondary = fig1_increase_species_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_increase_species_secondary = plot_clme(summ_fit2, 
                                            group = c("G1", "G2", "G3", "G4", "G5"),
                                            y_min = 0, y_max = 1.4, p_gap = 0.2, ann_pos = 1.3)
fig2_increase_species_secondary = fig2_increase_species_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

#### Monotonically decreasing trend test

```{r, fig.width=12, fig.height=5}
cons = list(order = "simple", decreasing = TRUE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_decrease_species_secondary = plot_clme(summ_fit1, 
                                            group = c("G1", "G2", "G3", "G4", "G5"),
                                            y_min = 0, y_max = 7, p_gap = 0.8, ann_pos = 6.5, decreasing = TRUE)
fig1_decrease_species_secondary = fig1_decrease_species_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_decrease_species_secondary = plot_clme(summ_fit2, 
                                            group = c("G1", "G2", "G3", "G4", "G5"),
                                            y_min = 0, y_max = 1.4, p_gap = 0.1, ann_pos = 1.3, decreasing = TRUE)
fig2_decrease_species_secondary = fig2_decrease_species_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

```{r, fig.height=10}
fig = ggarrange(fig1_increase_species_secondary, fig2_increase_species_secondary, 
                fig1_decrease_species_secondary, fig2_decrease_species_secondary, 
                nrow = 2, ncol = 2, common.legend = TRUE, labels = c("a", "", "b", ""))
fig
```

### Genus level {.tabset}

```{r}
d_alpha = alpha(genus_rarefied, index = c("observed", "diversity_shannon"))
df_alpha = data.frame(sampleid = meta(genus_rarefied)$sampleid,
                      subjid = meta(genus_rarefied)$subjid,
                      group = meta(genus_rarefied)$group2,
                      abx_use = meta(genus_rarefied)$abx_use,
                      Observed = d_alpha$observed,
                      Shannon = d_alpha$diversity_shannon,
                      check.names = FALSE) %>%
  filter(group != "missing")
df_alpha$group = recode(df_alpha$group, 
                        `g1` = "G1", `g2` = "G2",
                        `g3` = "G3", `g4` = "G4", `g5` = "G5")

df_fig = df_alpha %>% 
  gather(key = "measure", value = "value", Observed:Shannon)
df_fig$group = factor(df_fig$group, levels = c("G1", "G2", "G3", "G4", "G5"))
df_fig$measure = factor(df_fig$measure, 
                        levels = c("Observed", "Shannon"))
stat_test1 = df_fig %>%
  filter(measure == "Observed") %>%
  lm(formula = value ~ group + abx_use)
stat_test2 = df_fig %>%
  filter(measure == "Shannon") %>%
  lm(formula = value ~ group + abx_use)
```

#### Observed genus

```{r}
summ(stat_test1)
```

#### Shannon index

```{r}
summ(stat_test2)
```

#### Box plot

```{r, fig.width=10, fig.height=5}
df_stat = data.frame(p = round(summary(stat_test1)$coef[2:5, "Pr(>|t|)"], 2)) %>%
  mutate(x = "group", y = "value", measure = "Observed", 
         group1 = "G1", group2 = c("G2", "G3", "G4", "G5"),
         y.position = max(stat_test1$model$value) + seq(0, 3, 1)) %>%
  bind_rows(
    data.frame(p = round(summary(stat_test2)$coef[2:5, "Pr(>|t|)"], 2)) %>%
      mutate(x = "group", y = "value", measure = "Shannon", 
             group1 = "G1", group2 = c("G2", "G3", "G4", "G5"),
             y.position = max(stat_test2$model$value) + seq(0, 0.6, 0.2))
  ) 

p = df_fig %>%
  ggboxplot(x = "group", y = "value", color = "group",
            add = "jitter", 
            xlab = FALSE, ylab = FALSE, title = FALSE) +
  stat_pvalue_manual(df_stat, label = "p", tip.length = 0.01) +
  scale_color_npg(name = "Group") +
  facet_wrap(.~measure, scale = "free", nrow = 1) + 
  theme_bw() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5))
p
```

#### Monotonically increasing trend test

```{r, fig.width=12, fig.height=5}
df_alpha$group = factor(df_alpha$group, levels = c("G1", "G2", "G3", "G4", "G5"), ordered = TRUE)
cons = list(order = "simple", decreasing = FALSE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_increase_genus_secondary = plot_clme(summ_fit1, 
                                          group = c("G1", "G2", "G3", "G4", "G5"),
                                          y_min = 0, y_max = 12, p_gap = 0.8, ann_pos = 11)
fig1_increase_genus_secondary = fig1_increase_genus_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_increase_genus_secondary = plot_clme(summ_fit2, 
                                          group = c("G1", "G2", "G3", "G4", "G5"),
                                          y_min = 0, y_max = 2.5, p_gap = 0.2, ann_pos = 2.2)
fig2_increase_genus_secondary = fig2_increase_genus_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

#### Monotonically decreasing trend test

```{r, fig.width=12, fig.height=5}
cons = list(order = "simple", decreasing = TRUE, node = 1)

fit1 = clme(Observed ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit1 = summary(fit1)

fig1_decrease_genus_secondary = plot_clme(summ_fit1, 
                                          group = c("G1", "G2", "G3", "G4", "G5"),
                                          y_min = 0, y_max = 12.5, p_gap = 1.1, ann_pos = 11.5, decreasing = TRUE)
fig1_decrease_genus_secondary = fig1_decrease_genus_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Richness")

fit2 = clme(Shannon ~ group + abx_use, data = df_alpha, constraints = cons, seed = 123)
summ_fit2 = summary(fit2)

fig2_decrease_genus_secondary = plot_clme(summ_fit2, 
                                          group = c("G1", "G2", "G3", "G4", "G5"),
                                          y_min = 0, y_max = 2.5, p_gap = 0.2, ann_pos = 2.2, decreasing = TRUE)
fig2_decrease_genus_secondary = fig2_decrease_genus_secondary +
  scale_fill_npg(name = NULL) +
  labs(x = NULL, y = "Shannon Index")
```

```{r, fig.height=10}
fig = ggarrange(fig1_increase_genus_secondary, fig2_increase_genus_secondary, 
                fig1_decrease_genus_secondary, fig2_decrease_genus_secondary, 
                nrow = 2, ncol = 2, common.legend = TRUE, labels = c("c", "", "d", ""))
fig
```

## Beta diversity {.tabset}

1. P-values are obtained by PERMANOVA

2. Dissimilarity measure: Bray-Curtis

### Species level

```{r, fig.width=8, fig.height=6}
species_rarefied = subset_samples(species_rarefied, group2 != "missing")

dis = phyloseq::distance(species_rarefied, method = "bray")
mds  = ordinate(species_rarefied, "MDS", distance = dis)

# PCoA plot
fig_beta_species_secondary = plot_ordination(species_rarefied, 
                                             mds, color = "group2") +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(geom = "polygon", type = "t", level = 0.8,
               aes(fill = group2), alpha = 0.25) +
  scale_color_npg(name = NULL, labels = c("G1", "G2", "G3", "G4", "G5")) +
  scale_fill_npg(name = NULL, labels = c("G1", "G2", "G3", "G4", "G5")) +
  theme_bw()

# P-values
groups = meta(species_rarefied)$group2
set.seed(123)
stat_test = pairwise.perm.manova(resp = dis, fact = groups, nperm = 999, p.method = "none")
df_p = data.frame(X = rep("G1", 4),
                  Y = c("G2", "G3", "G4", "G5"),
                  p = round(stat_test$p.value[, "g1"], 2),
                  q = round(p.adjust(stat_test$p.value[, "g1"], method = "BH"), 2))

tab_beta_species_secondary = ggtexttable(df_p, rows = NULL, theme = ttheme("light"))

p = ggarrange(fig_beta_species_secondary, tab_beta_species_secondary,  
              ncol = 1, nrow = 2, heights = c(1, 0.4))
p
```

### Genus level

```{r, fig.width=8, fig.height=6}
genus_rarefied = subset_samples(genus_rarefied, group2 != "missing")

dis = phyloseq::distance(genus_rarefied, method = "bray")
mds  = ordinate(genus_rarefied, "MDS", distance = dis)

# PCoA plot
fig_beta_genus_secondary = plot_ordination(genus_rarefied, 
                                           mds, color = "group2") +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(geom = "polygon", type = "t", level = 0.8,
               aes(fill = group2), alpha = 0.25) +
  scale_color_npg(name = NULL, labels = c("G1", "G2", "G3", "G4", "G5")) +
  scale_fill_npg(name = NULL, labels = c("G1", "G2", "G3", "G4", "G5")) +
  theme_bw()

# P-values
groups = meta(genus_rarefied)$group2
set.seed(123)
stat_test = pairwise.perm.manova(dis, groups, nperm = 999, p.method = "none")
df_p = data.frame(X = rep("G1", 4),
                  Y = c("G2", "G3", "G4", "G5"),
                  p = round(stat_test$p.value[, "g1"], 2),
                  q = round(p.adjust(stat_test$p.value[, "g1"], method = "BH"), 2))

tab_beta_genus_secondary = ggtexttable(df_p, rows = NULL, theme = ttheme("light"))

p = ggarrange(fig_beta_genus_secondary, tab_beta_genus_secondary,  
              ncol = 1, nrow = 2, heights = c(1, 0.4))
p
```

# Outputs

```{r}
p_alpha = ggarrange(fig1_increase_species, fig2_increase_species, 
                    fig1_decrease_species, fig2_decrease_species, 
                    fig1_increase_genus, fig2_increase_genus, 
                    fig1_decrease_genus, fig2_decrease_genus,
                    ncol = 4, nrow = 2, common.legend = TRUE, legend = "top",
                    labels = c("a", "", "b", "", "c", "", "d", ""))
ggsave(filename = "../results/figures/supp_alpha.pdf", plot = p_alpha,
       width = 12, height = 8, units = "in")
ggsave(filename = "../results/figures/supp_alpha.jpeg", plot = p_alpha,
       width = 12, height = 8, units = "in", dpi = 100)

p_beta = ggarrange(fig_beta_species, tab_beta_species,  
                   fig_beta_genus, tab_beta_genus,  
                   ncol = 2, nrow = 2, widths = c(1, 0.4),
                   common.legend = TRUE, legend = "top",
                   labels = c("a", "", "b", ""))
ggsave(filename = "../results/figures/supp_beta.pdf", plot = p_beta,
       width = 10, height = 8, units = "in")
ggsave(filename = "../results/figures/supp_beta.jpeg", plot = p_beta,
       width = 10, height = 8, units = "in", dpi = 100)
```

# Session information

```{r, message = FALSE, warning = FALSE, comment = NA}
sessionInfo()
```
